The zoon R package for reproducible and shareable species distribution modelling

### 2017-04-10 response_to_reviewers.md

### Response to reviewers

Thank you very much for the rapid and detailed assessment of our manuscript. We'd like to thank the AE and reviewers for their positive comments, enthusiasm and very helpful suggestions, all of which have been taken on board in this revision. I have detailed our responses to these suggestions below, with text additions in boldface. I have also corrected all of the typos flagged during the review process (those corrections are not detailed here).

@@

Nick Golding, on behalf of the authors.

### #

---

### #

#### Reviewer 1 comments (email & pdf):

####

> "For example: add to the manuscript an attachment with the methods to build 'modules'."
>
> P10: "I just checked on the R Page, and it would be interesting to include as an attachment the basic steps to create the different modules that appear on that website."

@@

The zoon R package provides three vignettes, including a 4000-word vignette on how to build modules. We would rather users access this vignette via the package or the package's CRAN page than include it as a supplement, since that version can be updated along with future versions of the package (the vignette as at this release is also archived there). We have added the following sentence to the section *Example 2. Building a spatial thinning module* to refer explicitly to this vignette:

"**The zoon package vignette *Building modules* provides full details and examples of how to build modules of each type.**"

####

> "Finally, I want to make a mention to the platform to share data. The rights acquired by journals to publications that are not open access. This could generate some kind of legal conflict. It would be appropriate to mention something about the manuscript."
>
> P10: "This function is interesting and valuable for the scientific community. However, how do you handle the rights acquired by the Journal on the results they would be sharing on this platform?"

@@

#### Reviewer 2 comments:

### #

>  "First, there needs to be a clear definition of what a module is. For example, it wasn’t necessarily obvious that the modelling modules are not drawing directly from modelling functions already available (i.e. ‘randomForest’ vs ‘RandomForest’). Particularly, as people upload more modules with their own naming schemes, it may not be clear for less experienced users what the difference are and what the consequences of these differences could be for their analyses."

We have added the following clarifying text to the *Building a workflow* section, which introduces modules.

@@

"Each of these steps is encoded as one or more software modules **(snippets of R code) to complete one of these tasks. zoon modules can use methods from any R package, providing a common interface between the many SDM packages already available.**"

The potential issue of multiple similarly-named modules is closely related to Reviewer 1's comments above about filtering and exploring modules. In our edits in response to that comment, we discuss plans for a web platform where module developers and discuss, rate, and collaborate on modules. Ultimately, we expect the SDM community will be able to promote the best modules and ensure they are used, as happens already with R packages.

### #

@@

> "Finally, I think the authors may have missed an opportunity. An additional step in the workflow for testing if models are fit-for-purpose would be extremely beneficial i.e. testing model residuals forspatial autocorrelation, testing variables for excessive correlation, are the sample sizes large enough given the number of covariates, and are all the models in an ensemble useful. These are important considerations when fitting SDMs, and are far too often overlooked, particularly in packages such as ‘BioMod2’ and ‘sdm’. If we want to widen up the accessibility of SDMs beyond core modellers and improve their application by the less experienced, packages that aim to improve ease of use need to also offer opportunities for testing if a model has been fitted correctly. I understand that this will be a significant undertaking, but should be something that the authors consider, discuss, and potentially implement in a future version of the package."

We agree that automated checking and validation of analyses is an important and under-appreciated aspect of SDM modelling, and something that zoon could help to improve. Rather than adding an additional module type to zoon, such checks would be best handled via output modules. We as the developers of zoon do not want to be seen as dictating which methods are 'correct', but would prefer this be determined by consensus of the SDM community. We would therefore welcome the development by other parties of output modules that check whether workflows are fit for purpose. We have highlighted this opportunity in the Discussion, below an existing related paragraph:

"By providing a common interface for SDM and the capacity to create and share methodological experiments zoon offers a new opportunity for the SDM community to develop community modelling benchmarks, and resolve scientific and methodological questions in ways that the current culture of publishing cannot achieve.

**zoon also has the potential to improve methodological standards in SDM, a research area in which there is widespread misunderstanding and misuse of techniques (Yackulic *et al.* 2013). SDM methodologists could encode model validation procedures and best-practice guidance as zoon output modules, enabling end users to rapidly ensure their modelling procedures are fit for purpose (Guillera-Arroita *et al.* 2015).**"

### #

@@

> "Figure 1: What’s the difference between the grey and the black lines? Why in Figure 1C does the line come out of the bottom of the second process icon? Am I missing something?"

The colour and orientation of the arrows are stylistic and intended to focus the reader's eye on the relevant part of each diagram. Black lines are used to highlight the inputs and outputs of the module type which lists and chains are being applied. In Figure 1C the black line goes out of the bottom of the box for symmetry with the black line going into the box. We believe these choices help make this an engaging figure, and need not be detailed in the figure caption.

### 2017-04-13 ms.Rmd
add author contributions section

### Supporting Information

@@

>>>>

### Supporting information

@@

### Author's contributions
All authors contributed to the design of the zoon R package.
TAA, TCDL and NG implemented and programmed the R package with input from the other authors.
NG wrote the first draft of the manuscript and all authors contributed to subsequent revisions and gave final approval for publication.
